% ============================================================
\chapter{Assignments}
\label{chapter:assignments}
% ============================================================

This chapter addresses four of the seven optional assignments. Assignment~1 (combined network and functional architecture diagram) has already been completed in Chapter~\ref{chapter:infrastructure}, Section~\ref{sec:archimate-view}, using ArchiMate notation and TikZ. Assignment~2 maps business services to technical services and explains how this mapping could be represented in a monitoring dashboard. Assignment~3 designs a solution for status and performance monitoring of all servers. Assignment~7 designs a solution for logging and forensic data collection.

% ============================================================
\section{Assignment 2 -- Business Service to Technical Service Mapping}
\label{sec:assignment2}
% ============================================================

\subsection{Service Mapping}

An enterprise monitoring strategy must not only observe infrastructure components in isolation but must express the health of infrastructure in terms that are meaningful to the business. \label{tab:service-mapping}This requires an explicit mapping between business services, as understood by stakeholders, and the technical services and components that realise them.

The \textbf{Customer Order Portal} is realised by the Order System Cluster and depends on the full depth of the Eindhoven stack: the Perimeter Firewall, Load Balancer, Web Cluster, IAM/SSO server, Internal Firewall, Order System Cluster itself, MongoDB Cluster, RabbitMQ Cluster, and PostgreSQL Cluster. Because it is customer-facing and revenue-critical, any failure in this chain has direct business impact.

\textbf{Goods Storage and Distribution} is delivered across both sites. It depends on the Internal Firewall, Warehouse Cluster, RabbitMQ Cluster, and PostgreSQL Cluster in Eindhoven, as well as the WAN link (MPLS/VPN), the Groningen Perimeter Firewall, the WMS Node in Groningen, and the PostgreSQL Read Replica in Groningen. The WAN link is therefore a shared dependency between this service and the inter-site replication it requires.

\textbf{Financial Administration} relies on the Accounting Cluster, which in turn depends on the Internal Firewall, PostgreSQL Cluster, and the RabbitMQ Cluster for invoice creation events arriving asynchronously from the Order System. \textbf{Salary and Personnel Administration} follows a simpler dependency chain: the Internal Firewall, HR/Payroll Cluster, and PostgreSQL Cluster, with no message broker dependency.

\textbf{Identity and Access Management} is realised by the IAM/SSO server, which sits behind the Perimeter Firewall and uses the PostgreSQL Cluster as its user store. Its position as a shared dependency for every other service makes it uniquely critical: a single IAM failure blocks access to all six business services simultaneously. \textbf{IT Operations and Management} is supported by the Monitoring Server and Jump Host in the Management VLAN, both of which require IAM/SSO authentication for administrative access.

\subsection{Dependency Analysis}

The mapping reveals two critical shared dependencies that affect multiple business services simultaneously:

\begin{enumerate}
    \item \textbf{IAM / SSO Node:} A failure of the IAM/SSO node would render all six business services inaccessible, as every application requires authentication. This justifies its NFR-AV-06 availability target of 99.9\% and the recommendation to deploy it as a clustered pair.
    \item \textbf{PostgreSQL Cluster:} Four of the six business services depend on PostgreSQL. A total PostgreSQL failure would halt financial administration, HR/payroll, IAM authentication and the relational component of order processing. The hot-standby replica in Eindhoven and the automatic failover mechanism (via a connection pooler such as PgBouncer with health-check failover) are therefore non-negotiable.
\end{enumerate}

\subsection{Dashboard Design}

\subsubsection{Design Principles}

The monitoring dashboard is shaped by three principles drawn from the stakeholder analysis. The primary view must present service health in business terms --- ``Customer Order Portal: Operational'' rather than server metrics --- so that stakeholders without IT backgrounds, such as the Operations Director and CEO, can interpret it immediately. Role-based access then ensures that different stakeholders see the appropriate level of detail: business managers see service-level indicators, IT staff see the underlying component metrics, and no role sees data outside their responsibility. Finally, drill-down capability means that each business service indicator links to a second-level view of the contributing technical components, and a third level exposes raw metrics and logs for engineering investigation.

\subsubsection{Dashboard Structure}

The dashboard is structured across three tiers, each addressing a different audience.

\textbf{Tier 1 --- Executive / Business View (read-only, all authenticated users)} presents six status tiles, one per business service, each displaying the current operational state as \textcolor{green!60!black}{\textbf{Operational}}, \textcolor{orange}{\textbf{Degraded}}, or \textcolor{red}{\textbf{Unavailable}}. A 24-hour availability sparkline accompanies each tile, showing the percentage of time spent in the Operational state over the preceding day. If any P1 or P2 incident is active, a current incident banner is displayed at the top of the view.

\textbf{Tier 2 --- Operations View (IT staff and relevant department managers)} provides a per-service component health matrix showing the state, latency, and last check timestamp of each contributing infrastructure component. Key performance indicators are surfaced at this tier: the Order System's request rate and error rate, the Warehouse message queue depth, and database replication lag. WAN link status --- MPLS primary or VPN failover --- is shown with current bandwidth utilisation.

\textbf{Tier 3 --- Engineering View (IT staff only)} exposes raw infrastructure metrics. Individual server dashboards show CPU, memory, disk I/O, network throughput, and process count. Database dashboards cover query throughput, lock contention, replication lag in seconds, and the slow query log. RabbitMQ dashboards present queue depth per queue, consumer lag, and message publish and acknowledge rates. Firewall dashboards show connection table utilisation, blocked connection counts, and bandwidth per VLAN.

\subsubsection{Tooling Recommendation}

\textbf{Grafana} is recommended as the dashboard platform, consuming metrics from \textbf{Prometheus} (time-series metrics) and \textbf{Loki} (log aggregation). This stack is open-source, runs on the existing monitoring server in the Management VLAN, and supports role-based access control via LDAP/SAML integration with the IAM. Grafana's ``status panel'' plugin directly supports the traffic-light service health tiles described in Tier 1.

% ============================================================
\section{Assignment 3 -- Status and Performance Monitoring Solution}
\label{sec:assignment3}
% ============================================================

\subsection{Scope and Objectives}

This monitoring solution addresses status and performance monitoring for all server-class infrastructure components in the LogiFlux environment, across both sites. Derived from the requirements in Chapter~\ref{chapter:requirements}, the solution must provide real-time visibility into the operational state of all production servers and detect failures or degraded performance before they impact business services. It must support proactive capacity planning in line with NFR-CAP-01 through NFR-CAP-04, deliver automated alerting within two minutes of a critical failure as required by MR-02, and provide the availability data needed for monthly SLA reporting as required by MR-04.

\subsection{Architecture}

The monitoring architecture follows an \textbf{agent-based collection model}: a lightweight metrics agent runs on each monitored host and exposes data to a central Prometheus server in the Management VLAN. This avoids the need to open inbound firewall rules to production servers and reduces the overall attack surface.

For metrics collection, Prometheus uses \textbf{node\_exporter} on all Linux hosts and \textbf{windows\_exporter} on all Windows hosts, with additional service-specific exporters deployed per component type as detailed in Table~\ref{tab:exporters}. Prometheus retains 15 days of high-resolution metrics locally at a 15-second scrape interval; data is then downsampled and forwarded to \textbf{Thanos} or \textbf{VictoriaMetrics} for the 13-month long-term retention required by MR-06. Grafana, as described in Assignment~2, provides dashboards for all three stakeholder tiers. Alerting is handled by Prometheus \textbf{Alertmanager}, which routes alerts by severity: P1 and P2 alerts reach on-call IT staff via PagerDuty, while P3 and informational alerts are published to a Teams or Slack channel. The two-minute detection-to-notification requirement from MR-02 is met by a one-minute scrape interval on critical components combined with a one-minute alert evaluation cycle.

\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Component} & \textbf{Exporter / Agent} & \textbf{Key Metrics Collected} \\
\midrule
All Linux servers & node\_exporter & CPU (user/sys/iowait), memory (used/cached/available), disk I/O (read/write IOPS, latency), network (bytes in/out, errors), load average, uptime \\
All Windows servers & windows\_exporter & CPU utilisation, memory commit, disk queue length, network interface throughput, Windows event log counters, service state \\
PostgreSQL & postgres\_exporter & Connections (active/idle/waiting), transactions per second, replication lag (seconds), cache hit ratio, deadlocks, longest running query duration \\
MongoDB & mongodb\_exporter & Operations per second (reads/writes/commands), memory resident set, replication oplog window, connections, collection scan counts \\
RabbitMQ & rabbitmq\_exporter & Queue depth per queue, message publish/deliver/ack rates, consumer count per queue, unacknowledged messages, memory and disk alarms \\
Firewalls (pfSense/OPNsense) & pfsense\_exporter / SNMP & Connection table utilisation (\%), blocked packets/s, bandwidth per interface, CPU/memory of firewall appliance \\
WAN link & SNMP on WAN router & Interface bandwidth utilisation (\%), packet loss (\%), round-trip latency (ms) to Groningen, link state \\
\bottomrule
\end{tabularx}
\caption{Exporters and key metrics per infrastructure component}
\label{tab:exporters}
\end{table}

\subsection{Alert Thresholds}

Alert thresholds are defined at two levels. \textbf{P1} alerts require immediate response (on-call engineer notified within 2 minutes). \textbf{P2} alerts require same-day response. \textbf{P3} alerts are informational and visible in the dashboard only.

\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{lllX}
\toprule
\textbf{Metric} & \textbf{P1 (Critical)} & \textbf{P2 (Warning)} & \textbf{Notes} \\
\midrule
Server CPU utilisation & $>$90\% for 5 min & $>$70\% for 10 min & Applies to all production nodes \\
Server memory available & $<$5\% for 5 min & $<$15\% for 10 min & \\
Disk space remaining & $<$5\% & $<$15\% & Evaluated per-mount \\
Disk I/O latency & $>$50 ms avg & $>$20 ms avg & \\
Server unreachable & Immediate (node down) & --- & Exporter target disappears from Prometheus \\
PostgreSQL replication lag & $>$30 s & $>$10 s & Triggers before standby falls too far behind \\
PostgreSQL connections & $>$90\% of max & $>$75\% of max & \\
RabbitMQ queue depth & $>$10,000 messages & $>$5,000 messages & Per business-critical queue \\
WAN link utilisation & $>$90\% sustained & $>$70\% sustained & MPLS or VPN link \\
WAN packet loss & $>$5\% & $>$1\% & Over a 5-minute window \\
\bottomrule
\end{tabularx}
\caption{Alert thresholds for status and performance monitoring}
\label{tab:alert-thresholds}
\end{table}

\subsection{Groningen Site Monitoring}

The monitoring server in Eindhoven's Management VLAN collects metrics from Groningen components over the WAN link. Groningen exporters are scraped by the central Prometheus instance via the MPLS link. A \textbf{blackbox exporter} probe is run from Eindhoven to the Groningen firewall's management IP every 30 seconds to provide independent WAN liveness detection.

In the event of a WAN failure, Prometheus marks all Groningen targets as \textit{unreachable} and fires a P1 alert. If a local monitoring agent is desired at the Groningen site, a lightweight \textbf{Prometheus remote write agent} (Grafana Agent or Prometheus in agent mode) can buffer metrics locally and forward them when the WAN recovers.

\subsection{SLA Reporting}

Grafana's built-in \textbf{SLA/availability reporting} capability generates monthly reports from Prometheus availability data. The report for the Customer Order Portal (NFR-AV-01: 99.9\%) is calculated as:

\[
\text{Availability} = \frac{\text{Total minutes in month} - \text{Minutes in P1 outage state}}{\text{Total minutes in month}} \times 100\%
\]

These reports are exported as PDF and sent automatically to the IT Director and Operations Director at the end of each calendar month (MR-04).

% ============================================================
\section{Assignment 7 -- Logging and Forensic Data Collection}
\label{sec:assignment7}
% ============================================================

\subsection{Objectives}

A centralised logging and forensic readiness solution serves three distinct purposes at LogiFlux:

\begin{enumerate}
    \item \textbf{Operational logging:} Capturing application and system events to support incident diagnosis, troubleshooting and root-cause analysis.
    \item \textbf{Compliance logging:} Retaining audit trails required by Dutch law (fiscal records: 7 years), GDPR (access logs and data processing records) and internal policy.
    \item \textbf{Forensic readiness:} Ensuring that in the event of a security incident, sufficient evidence is available in an admissible, tamper-evident form to support investigation and legal proceedings.
\end{enumerate}

\subsection{Log Sources}

\label{tab:log-sources}The LogiFlux environment generates log data from twelve distinct sources, grouped here by category.

Security and network logs originate from the perimeter firewalls at both Eindhoven and Groningen --- capturing connection allow/deny decisions, NAT translations, IDS/IPS alerts, and interface state changes --- and from the internal firewall at Eindhoven, which records inter-VLAN traffic decisions and policy violation events. The IAM/SSO server is also a critical security log source, producing authentication events (success and failure), MFA events, session creation and destruction records, and role assignment changes. Windows endpoints at both sites contribute Windows Security Event Log entries covering logon/logoff, privilege use, and process creation (Event IDs 4624, 4625, 4648, 4688, and 4697). The jump host generates the most forensically significant logs: full session recordings capturing keystrokes and timing for all administrative sessions via Teleport or equivalent, alongside command history.

Operational logs come from the load balancer (HTTP access log including method, URI, response code, client IP, and response time, plus backend health check results), the web cluster (application access logs, HTTP error logs, and SSL handshake errors), the RabbitMQ cluster (message publish, consume, ack, and nack events; queue creation and deletion; consumer connect and disconnect), MongoDB (slow query log for queries exceeding 100~ms, authentication events, and replication events), and the monitoring server itself (Prometheus scrape errors and Alertmanager notification delivery and failure records).

Compliance logs are generated by the application clusters --- all four produce application error logs and inter-service API call logs, while the Accounting and HR clusters additionally produce user action audit logs attributing each operation to a named identity --- and by PostgreSQL, which logs all DDL operations, DML operations on sensitive tables, connection events, authentication failures, and vacuum/checkpoint events.

\subsection{Centralised Log Architecture}

All log sources forward data to a centralised log management platform running on the monitoring server in the Management VLAN. \textbf{Fluent Bit} serves as the lightweight forwarder, running as a daemon on all hosts; it collects, parses, and forwards logs over TLS to the central aggregator. Fluent Bit is chosen over Filebeat for its lower resource footprint, which is particularly relevant on the resource-constrained WMS node at Groningen. \textbf{Loki} handles log aggregation, storing logs indexed by labels (host, service, VLAN, and site); its label-based indexing model is highly efficient for the LogiFlux log volume and integrates natively with Grafana for unified log and metrics exploration. IT staff query logs using \textbf{Grafana}'s Explore interface with LogQL --- the same platform used for performance dashboards in Assignment~2 --- avoiding the need to deploy a separate UI for logs.

For long-term cold storage (compliance retention beyond 30 days), Loki is configured with an \textbf{S3-compatible object storage backend} (on-premises MinIO instance or a compliant cloud bucket). This separates hot operational logs (fast SSD storage, 30-day retention) from cold compliance logs (object storage, 7-year retention for financial audit logs, 36-month retention for security logs per NFR-SEC-05).

\subsection{Retention Policy}

\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{lllX}
\toprule
\textbf{Log Category} & \textbf{Hot (fast)} & \textbf{Cold (object)} & \textbf{Legal/Policy Basis} \\
\midrule
Security logs (FW, IAM, Jump Host) & 30 days & 36 months & NFR-SEC-05 / GDPR / incident response \\
Application audit logs (Accounting) & 90 days & 7 years & Dutch fiscal law (Belastingdienst) \\
Application audit logs (HR/Payroll) & 90 days & 5 years & GDPR / Dutch employment law \\
Operational logs (web, app servers) & 14 days & 6 months & Internal policy \\
RabbitMQ / DB operational logs & 14 days & 3 months & Internal policy \\
Jump Host session recordings & 90 days & 24 months & Security policy / forensic readiness \\
Windows endpoint event logs & 14 days & 12 months & Security policy \\
\bottomrule
\end{tabularx}
\caption{Log retention policy by category}
\label{tab:retention}
\end{table}

\subsection{Forensic Readiness}

Forensic readiness means that the organisation can respond to a security incident or legal request without having to begin evidence collection after the fact. The following measures are implemented:

\begin{enumerate}
    \item \textbf{Tamper-evident log shipping:} Logs are forwarded from source hosts to the central Loki instance via TLS-encrypted, authenticated channels (mTLS between Fluent Bit agents and the Loki ingestor). Once stored in object storage, logs are write-once (object lock / immutability policy) to prevent post-incident modification.

    \item \textbf{Jump Host session recording:} All administrative sessions through the jump host are recorded using \textbf{Teleport} (or equivalent PAM solution), producing a timestamped, replay-capable session log. These recordings are stored in the object store under an immutable retention policy and are immediately available for investigation. This directly satisfies MR-05 and UR-05.

    \item \textbf{Log integrity:} Each batch of logs shipped to the central aggregator includes an HMAC checksum, allowing verification that no log entries have been deleted or altered in transit or storage.

    \item \textbf{Chain of custody documentation:} A documented procedure defines who may access log data for forensic purposes, under what conditions, and how evidence is preserved and handed over to external parties (law enforcement, auditors). This procedure is owned by the IT Director.

    \item \textbf{Forensic collection triggers:} If a P1 security alert is fired, Alertmanager automatically triggers a log snapshot job that archives a 72-hour log window from all sources into a cryptographically signed ZIP archive in isolated storage. This preserves evidence even if an attacker subsequently attempts to delete log data.
\end{enumerate}

\subsection{Information Flow}

The logging information flow is as follows:

\begin{enumerate}
    \item Sources generate log events (application logs, syslog, Windows Event Log, session recordings).
    \item Fluent Bit agents on each host parse, structure (JSON) and enrich events with metadata (hostname, VLAN, site, service name).
    \item Events are forwarded over TLS to the Loki ingestor on the Monitoring Server.
    \item Loki stores hot logs on local SSD. A background job streams logs older than the hot-tier threshold to the MinIO object store.
    \item Grafana provides query access to all tiers via LogQL. Role-based access control limits which log streams each user may query.
    \item Alertmanager monitors Loki for pattern-based log alerts (e.g. repeated authentication failures, firewall rule violations) and notifies IT staff via the same alert routing defined in Assignment~3.
\end{enumerate}

\subsection{GDPR Considerations}

Centralised logging inevitably involves processing personal data (usernames, IP addresses, potentially employee activity records). Log data is pseudonymised where possible, with internal user identifiers replacing full names in non-audit logs. Access to logs containing personal data --- HR audit logs and employee endpoint logs --- is restricted to the IT Director and HR Manager through Grafana's RBAC policy. A \textbf{Data Processing Register} entry is maintained for the logging system, documenting data categories, retention periods, and access controls as required by GDPR Article~30. All logs are stored within the EU, on-premises in the Netherlands, with any cloud object storage restricted to an EU region.
